A Sample {\ttlit ACM} SIG Proceedings Paper in LaTeX Format\titlenote{(Does NOT produce the permission block, copyright information nor page numbering). For use with ACM\_PROC\_ARTICLE-SP.CLS. Supported by ACM.}

### 2014-03-07 paper/draft.tex
First revision

%%% PAGE DIMENSIONS
%\usepackage{geometry} % to change the page dimensions
%\geometry{a4paper} % or letterpaper (US) or a5paper or....
% \geometry{margin=2in} % for example, change the margins to 2 inches all round
% \geometry{landscape} % set up the page for landscape
%   read geometry.pdf for detailed page layout information
\usepackage[hidelinks]{hyperref}
\usepackage[autostyle,french=guillemets,babel]{csquotes}
\usepackage{graphicx} % support the \includegraphics command and options
\usepackage[backend=biber]{biblatex}
\addbibresource{biblio.bib}
% \usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent

%%% PACKAGES
\usepackage{booktabs} % for much better looking tables
\usepackage{array} %
% for better arrays (eg matrices) in maths
\usepackage{paralist} % very flexible & customisable lists (eg. enumerate/itemize, etc.)
\usepackage{verbatim} % adds environment for commenting out blocks of text & for better verbatim
\usepackage{subfig} % make it possible to include more than one captioned figure/table in a single float
% These packages are all incorporated in the memoir class to one degree or another...
\usepackage{amsfonts}

@@

\title{A Sample {\ttlit ACM} SIG Proceedings Paper in LaTeX
Format\titlenote{(Does NOT produce the permission block, copyright information nor page numbering). For use with ACM\_PROC\_ARTICLE-SP.CLS. Supported by ACM.}}
\subtitle{[Extended Abstract]
\titlenote{A full version of this paper is available as
\textit{Author's Guide to Preparing ACM SIG Proceedings Using
\LaTeX$2_\epsilon$\ and BibTeX} at
\texttt{www.acm.org/eaddress.htm}}}
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\numberofauthors{1}
\alignauthor
Ben Trovato\titlenote{Dr.~Trovato insisted his name be first.}\\
       \affaddr{Institute for Clarity in Documentation}\\
       \affaddr{1932 Wallamaloo Lane}\\
       \affaddr{Wallamaloo, New Zealand}\\
       \email{trovato@corporation.com}
% 2nd. author
%\date{} % Activate to display a given date or no date (if empty),
         % otherwise the current date is printed 
}
\begin{document}
\maketitle

@@

The second class of issues concerns platform integrity and user data privacy. On one hand, from an user-centric point of view, the user would like to preserve privacy of its data when it stores data on the cloud platform. However, hypervisor is normally the most privileged task executed in the platform and therefore a curious or malicious administrator could easily retrieve those data \cite{cloudvisor:zhang}.
On the other hand, from a user-centric point of view, it is difficult to observe and inspect deeply the functioning of the whole platform, leveraging commodity general purpose hypervisors (GPH).General Purpose Hypervisors are executed at the most privileged level and so, it is complicated to detect and intervene on them in presence of malicious attack or misbehavior, without service interruption for all VMs hosted on the server. 
To sum up, both visions collide with the execution of the hypervisor at the highest level of privileged. This attitude prevents the platform administrator from a real monitoring of the whole platform, but also the possibility for users to concretely inspects actions that platform administrators do on user data. In the following paragraphs, we will make reference to this class of problem as ``Vertical'' requirements.

To resume, in order to present a complete view of the architecture, we identified five key design principles, that a cloud computing platform has to implement in order to solve the cited problems. In the following paragraphs, those principles will represent a grid of analysis of different architectures.

\paragraph{Inter-platform support} The cloud architecture has to be designed in order to support directly different platforms and service providers. 
\paragraph{Fail-Safe modular architecture} In order to recover soft failures, the archtecure is supposed to be fail-safe, being able of automatically migrate instances or user-data in presence of a menace, but also reactivating the correct functioning in presence of a software failure. In order to satisfy this property, the amount of safety-critical components has to be as lowest as possible and the overall design has to be kept as modular as possible
\paragraph{Small TCB} The amount of lines of code executed at the highest level of privilege is normally proportional to the number of potential failures and bugs in the platform. According to this, an architecture presenting a tiny TCB will improve its safety and integrity properties by design.
\paragraph{User-Data privacy} A curious or malicious administrator ( or another platform user) would not have the right to inspect and analyze user data and instances, without the explicit consensus with the user above mutually trusted services. User-data privacy is a major challenge of modern cloud platforms. It is important to identify a trade-off that allows providers to identify and intervene on possible menaces, but also guarantee the impossibility for a curious or malicious operator to access to them without restrictions.
\paragraph{3rd party compatibility} Avoid disruption of existing 3rd part applications compatibility could represent a major principle to consider. Platform interfaces has to be kept as close as possible to the existing one, preventing application rewrite.

Moreover, beside this functional design principles specifying which features have to be achieved by the platform, another set of non-design principles could be identified. Those principles does not concern directly features or aspects of the architecture of the platform, but the effective feasibility of the architecture. 
\paragraph{Minimize code addition} a major issue could be represented by the amount of code lines that have to be added and maintained to an existing architecture to meet the functional requirement. Even it does not represent strictly a technical limitation to a certain proposal, it could become a ``show-stopper'' factor for its utilization.

In our analysis, it will be kept in consideration not only the overall suitability of the architecture, but also a quantification of the effort needed to develop and deploy it. In table \ref{int:des} it is possible to observe a resume of previous cited principles.

\begin{table}
\label{int:des}
\begin{tabular}{llll}
\toprule
Id & Functional design Principle & Class &\\
\midrule
   DS1 & User-Data privacy & Vertical & \\
   DS2 & Inter-platform support & Horizontal & \\
   DS3 & Small TCB & Vertical & \\
   DS4 & Fail-Safe modular architecture & Vertical & \\ 
   DS5 & Backwards 3rd party compatibility & Horizontal \\
   NDS1 & Minimize code addition & $\times$& \\
\bottomrule
      \end{tabular}
\end{table}

@@

To sum up, no architecture had been presented to address vertical and horizontal class of problems at the same time until now. XenBlanket does not investigate cloud architectural security concerns (DS3,DS4), limiting the focus on new potential feature of a coordinated multi-provider cloud. Cloudvisor proposed a security architecture that leverages nested virtualization without allowing the possibility of executing more than one nested hypervisor at time, that is a crucial requirements of a multi-provider cloud platform (DS2). In a SuperCloud scenario, each user has to run its own L1 hypervisor instance to enforce a compatibility layer, overcoming incompatibilities between different providers.

Other approaches leveraging only one layer of virtualization does not achieve the whole design-principle satisfaction. Self-service cloud approach (SSC) \cite{ssc:art} design a complete security architecture to overcome major security and privacy issues of currently deployed cloud infrastructures, in order to alleviate the different needing user-centric and provider-centric vision of cloud computing presented in \ref{par:intro}. Even if this architecture reproduces an approach similar to the SuperCloud without leveraging an extra layer of virtualization ( DS

@@

Secondly, due to the monolithic architecture, each feature introduced in GPH enlarges the TCB. Therefore, leveraging GPH, a huge amount of code is running in the most privileged mode and has to be trust. Several works discuss and analyze the overall TCB of modern GPH \cite{nova, xmhf}, considering GPH TCB always bigger than 100 KLoC. This approach directly collides with one design principles introduced in the introduction paragraph (DS3), that identified a small TCB as a major requirement for a solid architecture.
%A brief comparison of TCB of widespread GPH is depicted in figure \ref{fig:tcbchart}. 
%More importanly, this is particularly true looking at the monolithic approach of commercial hyper-visors, where any addition to the hypervisor body enlarges the TCB.

Finally, the above architecture does not improve consistently the overall stability of the infrastructure. The underlying core, L0, has the same problem of single-layer GPH architecture. Infact, the consistent TCB implies also that a failure experienced in one of the critical components will simply produce a generalized failure of the whole platform, not satysfing the DS4. This kind of architecture, in particular, does not leverage completely the possibility to hide behind a compatible layer, a really modular architecture with only a tiny critical core.

To sum up, current nested virtualization limits are represented by a certain lack of dedicated feature in modern hypervisor and the impossibility to rely on a tiny TCB (DS3,DS4).
The first one is being addressed by hypervisor developers. The second one could simply not be achieved with modern hypervisors.

@@

Therefore, leveraging this concept of microkernels, a micro hypervisor (MH) is composed from a small kernel providing only essential functions, like scheduling, IPC, memory management and VMX extension. Several micro-hypervisor have been proposed in literature \cite{xmhf, nova}. They rely on the micro-kernel paradigm, extending it to virtualization. MH naturally satisfy vertical design principles (DS1,DS3, DS4). The tiny core could be considered the only system-critical componente of the architecture and if it possible to assure its correctness it could represent the last line of defense, where it could be possible to rebuild the architecture.

@@

\subsection {MH open issues}

@@

Moreover, as reported in previous sections, several cloud computing issues are simply not addressable with general purpose hypervisor (GPH). However, they are commonly widely adopted and represent the bedrock of existing platform. Therefore, GPH exports wide and solid interfaces leveraged by a plenty of different software that is able to properly communicate and act through them on the virtualization layer. On the contrary, MH lack of cloud toolkit integration, requiring an explicit support to be written, not satysfing 3rd party compatibility requirements (DS5). GPH are tightly connected with cloud toolkit and could present solid interfaces to communicate with different tools.

Finally, for sake of keeping the TCB small, MH could generally lack of several virtualization features, commonly available on modern GPH hypervisor. For instance, XMHF does not support multiple guests. This kind of drawbacks could represent important issues to handle and limit the potential interest in MH, even they are able to introduce several important features.

@@

The first analyzed architecture concerns only GPH. At first, it is important to recall the most interesting achievable by such architectures. In particular, several remarkable features are live migration over different platform, transparent migration of virtualized clusters between different data-centers and improved isolation between customer instances and administration, improved control of the nested virtualization layer L1 (horizontal requirements DS.
This approach allows us also to exploit different virtualization techniques. For instance, it could be possible to run at the same time PV and HVM nested instances, leveraging  Xen on KVM architecture ( if KVM nested VMX performs as expected).
The choice of KVM as L0 is due to the general smaller TCB, compared with standard Xen (without stub domains) and this is important due to the high privilege level of L0.

Supposing the completeness and the correctness of nested virtualization hardware facilities in widespread hypervisor, GPH over GPH represents the best achievable architecture leveraging mature hypervisors. To sum up, it could be possible to achieve live-migration between different platform and other features related to VM management flexibility but it could not still be possible to reduce the TCB. Analyzing MH approaches, the aim is try at most to retain to features, trying to introduce isolation and privacy improvements due to the lower TCB promised.

@@

MH on MH in reported only for completeness. The interest on such architecture is limited because it will still present all cited issues of MH.
Moreover, the lack of nested VMX functionalities prevents us to concretely test this kind of architecture, because MH normally rely on VMX facilities in order to create HVM guests.
To conclude, nested virtualization is exploited in order to fill eco-systems gaps with GPH but MH over MH does not propose any improvement in this sense.

@@

The second architecture, proposed in order to reduce security concerns, is built combining an GPH at L0 and a MH at L1. 
%Several MH are proposed in literature focusing on different .
%Therefore, this model tries to exploit microhypervisor as L1 hypervisor in order to enforce better isolation between L2 and L0. L1 will exploit the emulated nested VMX provided by GPH L0 and virtualize L2. 
%With this approach, we try to isolate the attacker in a VM virtualized on a secured layer, in order to prevent attacks to L0. The attacker would have to face with an essential virtualization layer and it would not be able to attack the back-end virtualization layer. 
An interesting point of this approach is the fact of being easily be deployed, without requiring any extra code addition to the GPH, L0.  Existing GPH-based platforms could potentially exports nested VMX features to those guests. Even if this is not so usual in modern cloud platform, this lack is considered temporary and related only to the immaturity of those features.
Secondly, the number of driver needed to properly support devices is significantly small. With a MH, the only I/O communication that the platform could be able to deliver, without the porting or addition of code, is Multi-level device assignment. This kind of approach is feasible only implementing in the L0 hypervisor software techniques like vIOMMU, that extends the IOMMU capabilities to several layers of virtualization.

In addition, the amount of code running in L0 kernel space is still as big as before and, therefore, no TCB reduction. 
Finally, adopting a MH as L1 prevents to use inter-platform enhanced features. As analyzed when dealing with MH design, the price of having a tiny core is normally paid with a lack of certain functionalities. For example, XMHF only supports one single guest. It could be easily noticed that this approach is not architecturally able to provide any new "horizontal" features, leveraging on nested virtualization. Therefore, the architecture is only able to provide the same features that are provided by today deployed ones, with an enhanced but not sufficient level of isolation.

@@

\begin{table*}
\centering
\begin{tabular}{llllll}
\toprule
Id & Functional design Principle & GPH over GPH & MH over MH & MH over GPH & GPH over MH\\
\midrule
DS1 & User-Data privacy & $\times$  & $\times$ & $\times$ & $\times$ \\
   DS2 & Inter-platform support & $\times$  & $\times$ & $\times$ & $\times$ \\
   DS3 & Small TCB & $\times$  & $\times$ & $\times$ & $\times$ \\
   DS4 & Fail-Safe modular architecture & $\times$  & $\times$ & $\times$ & $\times$ \\
   DS5 & Backwards 3rd party compatibility & $\times$  & $\times$ & $\times$ & $\times$ \\
   NDS1 & Minimize code addition & $\times$  & $\times$ & $\times$ & $\times$ \\
\bottomrule
      \end{tabular}
\end{table*}

In this configuration, the amount of code that have to be trusted is lower than other architectures, due to the modular structure of micro-kernels., executed at L0.
The possibility to exploit at L1 a GPH like Xen or KVM allows to deploy a similar architecture used in production today. The L1 GPH is normally integrated with modern cloud toolkits like OpenStack.
Therefore, the architecture presents a stronger isolation leveraging on nested virtualization, providing in particular a strong tiny core. The MH layer will be transparent from the user and it will not not any change in the functioning of the platform.

The first class of possible drawbacks of this architecture is represented by the necessity to expose to the L1 nested hypervisor the nested VMX extension.  This is experimentally done by GPH like Xen and KVM, but it is generally not included in the essential features provided by a minimal micro-hypervisor. As reported in XMHF paper \cite{xmhf}, the amount of lines of code for nested VMX is around 1KLOC. 
More generally, from an architectural point of view, the general MH minimalist attitude could represent a limitation to general preservation of existing platform functions. The constraint of privileged code size forces normally MH developers to drop some features, considered not essential. Therefore, several limitations could arise not only for nested virtualization proper features, but also for other class of features. For instance, XHMF is able only to virtualize one L1 instance at time. This is acceptable if the platform objective is limited to be provided of a secure core, but dramatically limits the possibility of having a platform model able to give the user the possibility to run its proper hypervisor as L1.

@@

To conclude, this section provides a short comparison between different platforms, contrasting architecture features above the two main different types of benefits given by nested virtualization, as reported widely in literature \cite{art:blan, nested:berghmans,  cloudvisor:zhang, turtle:ibm}. The first could be identified in the augmented degree of isolation the architecture could enforce between host OS and the "customer" point of access, leveraging a trusted core. The second one is represented by the possibility to leverage multiple virtualization layers to easily define inter-platform policy and operations between different platforms.
In the above paragraphs, four different possible architectures were presented with the set of  substitute a GPH  with a MH, looking at the positive and negative new features. The starting point is GPH over GPH, a well analyzed approach in literature \cite{turtle:ibm, art:blan, }. We underlined features and limitations of this architecture in the first paragraphs and in the

To conclude, GPH over MH is the most interesting one in order to enlarge nested virtualization promised benefit. This architecture with a solid and secured MH as L0 is the only able to reduce the platform TCB and this could be a key point to obtain more protected infrastructure, .
Firstly, adopting a MH as architecture trusted core, we will have a secure layer guaranteeing that administrator could not lose the control of the system. Klein et Al. \cite{Klein:2009} have demonstrated that a kernel with less than 10KLoC could be formally verified, under several conditions. This is far from be achievable today with standard GPH over GPH approach where the core is not tiny and the TCB is an order of 
magnitude bigger than MH \cite{nova}.

However, the effort to set up a GPH over MH is strongly dependent on which technique leveraging for virtualizing L2. As stated in paragraph \ref{par:gom}, with L2 guest virtualized with the hardware assitance technique, the most popular one, the MH needs to export nVMX at least. 
On the contrary, adopting paravirtualization/DBT for L2 does not necessitate a particular hardware support and so it could not be necessary patch the MH to set up a platform. 
However, it has not be underestimated, however, that para-virtualization supports only a subset of OS, and could not be considered a general solution and DBT has shown important overhead.
%The last point of interest is related to device management. If MH act as L0 hypervisor, it has to deal with hardware devices, and this could necessitate of a significant code writing. In fact, there are several technologies, like Direct Assignment, allowing to bypass hypervisor control of a device. Those techniques are useful but they reduce the flexibility in the device management, requiring for example exclusive allocation of a device to a guest. To contrast this drawbacks, PCI-E propose the possibity to multiplex device with a set of virtual devices called virtual functions proposed to the L0 hypervisor, that will relax the contraint of exclusive allocation. Unfortunately this approach could not give the same flexibilty given by other techniques, like para-virtualization 
%With MH over GPH, MH could directly assigne devices to GPH that is able to implement  para-virtualize device, for example. 
%This will relief the MH of a part of devices management, giving to the GPH the responsibility to drive concretely the device but without losing interesting features. Therefore, we will need a limited amount of code to handle device without have to deal with each specific device.

>>>>

%%% PAGE DIMENSIONS
%\usepackage{geometry} % to change the page dimensions
%\geometry{a4paper} % or letterpaper (US) or a5paper or....
% \geometry{margin=2in} % for example, change the margins to 2 inches all round
% \geometry{landscape} % set up the page for landscape
%   read geometry.pdf for detailed page layout information
\usepackage[hidelinks]{hyperref}
\usepackage[autostyle,french=guillemets,babel]{csquotes}
\usepackage{graphicx} % support the \includegraphics command and options
\usepackage[backend=biber]{biblatex}
\addbibresource{biblio.bib}
% \usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent
%%% PACKAGES
\usepackage{booktabs} % for much better looking tables
\usepackage{array} %
% for better arrays (eg matrices) in maths
\usepackage{paralist} % very flexible & customisable lists (eg. enumerate/itemize, etc.)
\usepackage{verbatim} % adds environment for commenting out blocks of text & for better verbatim
\usepackage{subfig} % make it possible to include more than one captioned figure/table in a single float
% These packages are all incorporated in the memoir class to one degree or another...
\usepackage{amsfonts}

@@

\title{A Sample {\ttlit ACM} SIG Proceedings Paper in LaTeX
Format\titlenote{(Does NOT produce the permission block, copyright information nor page numbering). For use with ACM\_PROC\_ARTICLE-SP.CLS. Supported by ACM.}}
\subtitle{[Extended Abstract]
\titlenote{A full version of this paper is available as
\textit{Author's Guide to Preparing ACM SIG Proceedings Using
\LaTeX$2_\epsilon$\ and BibTeX} at
\texttt{www.acm.org/eaddress.htm}}}
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\numberofauthors{1}
\alignauthor
Ben Trovato\titlenote{Dr.~Trovato insisted his name be first.}\\
       \affaddr{Institute for Clarity in Documentation}\\
       \affaddr{1932 Wallamaloo Lane}\\
       \affaddr{Wallamaloo, New Zealand}\\
       \email{trovato@corporation.com}
% 2nd. author
%\date{} % Activate to display a given date or no date (if empty),
         % otherwise the current date is printed 
}

%\setlist[itemize]{leftmargin=*}
\begin{document}
\maketitle

@@

The second class of issues concerns platform integrity and user data privacy. On one hand, from an user-centric point of view, the user would like to preserve privacy of its data when it stores data on the cloud platform. However, hypervisor is normally the most privileged task executed in the platform and therefore a curious or malicious administrator could easily retrieve those data \cite{cloudvisor:zhang}. Moreover, in multi-tenancy cloud platform, a user shares the same physical resources with other users. Several works \ref{sec:you,bot:att} showed the potential threat and risks of attacks issued by other users located on the same physical resource.

On the other hand, from a user-centric point of view, it is difficult to observe and inspect deeply the functioning of the whole platform, leveraging commodity general purpose hypervisors (GPH).General Purpose Hypervisors are executed at the most privileged level and so, it is complicated to detect and intervene on them in presence of malicious attack or misbehavior, without service interruption for all VMs hosted on the server. 
To sum up, both visions collide with the execution of the hypervisor at the highest level of privileged. This attitude prevents the platform administrator from a real monitoring of the whole platform, but also the possibility for users to concretely inspects actions that platform administrators do on user data. In the following paragraphs, we will make reference to this class of problem as ``Vertical'' requirements.

To resume, in order to present a complete view of the architecture, we identified five key design principles, that a cloud computing platform has to implement in order to solve the cited problems. In the following paragraphs, those principles will represent a grid of analysis of different architectures.
\begin{itemize}[]
\item \textbf{Inter-platform support} The cloud architecture has to be designed in order to support directly different platforms and service providers. The same user virtual instances have to be executed ( and potentially migrated) unmodified over different platforms.
\item \textbf{Fail-Safe modular architecture} In order to recover soft failures, the archtecure is supposed to be fail-safe, being able of automatically migrate instances or user-data in presence of a menace, but also reactivating the correct functioning in presence of a software failure. In order to satisfy this property, the amount of safety-critical components has to be as lowest as possible and the overall design has to be kept as modular as possible
\item \textbf{Small TCB} The amount of lines of code executed at the highest level of privilege is normally proportional to the number of potential failures and bugs in the platform. According to this, an architecture presenting a tiny TCB will improve its safety and integrity properties by design.
\item \textbf{User-Data privacy} A curious or malicious administrator ( or another platform user) would not have the right to inspect and analyze user data and instances, without the explicit consensus with the user above mutually trusted services. User-data privacy is a major challenge of modern cloud platforms. It is important to identify a trade-off that allows providers to identify and intervene on possible menaces, but also guarantee the impossibility for a curious or malicious operator to access to them without restrictions. In order to respect this principle when a platform administrator manages the platform it has not the highest level of privilege and each action on users object that he wants to deliver has to be validated by a mutually-trusted core, executed at the highest level of privilege. Moreover, this mutually-trust core has also to apply memory page cipher/uncipher in order to prevent memory page snooping when a page is swapped out.
\item \textbf{3rd part compatibility} Avoid disruption of existing 3rd part applications compatibility could represent a major principle to consider. Platform interfaces has to be kept as close as possible to the existing one, preventing application rewrite.
\end{itemize}
Moreover, beside this functional design principles specifying which features have to be achieved by the platform, another set of non-design principles could be identified. Those principles does not concern directly features or aspects of the architecture of the platform, but the effective feasibility of the architecture. 
\begin{itemize}[]
\item \textbf{Minimize code addition} a major issue could be represented by the amount of code lines that have to be added and maintained to an existing architecture to meet the functional requirement. Even it does not represent strictly a technical limitation to a certain proposal, it could become a ``show-stopper'' factor for its utilization.
\end{itemize}
In our analysis, it will be kept in consideration not only the overall suitability of the architecture, but also a quantification of the effort needed to develop and deploy it. In table \ref{int:des} it is possible to observe a resume of previous cited principles.

\begin{table}
\label{int:des}
\begin{tabular}{llll}
\toprule
Id & Functional design Principle & Class &\\
\midrule
   DP1 & User-Data privacy & Vertical & \\
   DP2 & Fail-Safe modular architecture & Vertical & \\ 
   DP3 & Small TCB & Vertical & \\
   DP4 & Inter-platform support & Horizontal & \\
   DP5 & Backwards 3rd part compatibility & Horizontal \\
   NDP1 & Minimize code addition & & \\
\bottomrule
      \end{tabular}
\end{table}

@@

To sum up, no architecture had been presented to address vertical and horizontal class of problems at the same time until now. XenBlanket does not investigate cloud architectural security concerns (DP3,DP2), limiting the focus on new potential feature of a coordinated multi-provider cloud. Cloudvisor proposed a security architecture that leverages nested virtualization without allowing the possibility of executing more than one nested hypervisor at time, that is a crucial requirements of a multi-provider cloud platform (DP4). In a SuperCloud scenario, each user has to run its own L1 hypervisor instance to enforce a compatibility layer, overcoming incompatibilities between different providers.

Other approaches leveraging only one layer of virtualization does not achieve the whole design-principle satisfaction. Self-service cloud approach (SSC) \cite{ssc:art} design a complete security architecture to overcome major security and privacy issues of currently deployed cloud infrastructures, in order to alleviate the different needing user-centric and provider-centric vision of cloud computing presented in \ref{par:intro}. Even if this architecture reproduces an approach similar to the SuperCloud without leveraging an extra layer of virtualization (DP4), the approach changes the behavior where the user acts introducing new components that have to be directly managed by the user. Therefore, the main drawback of this solution is to change the interface with the user, preventing him to straightforwardly move to the new platform without having to change its applications (DP5).

@@

Secondly, due to the monolithic architecture, each feature introduced in GPH enlarges the TCB. Therefore, leveraging GPH, a huge amount of code is running in the most privileged mode and has to be trust. Several works discuss and analyze the overall TCB of modern GPH \cite{nova, xmhf}, considering GPH TCB always bigger than 100 KLoC. This approach directly collides with one design principles introduced in the introduction paragraph (DP3), that identified a small TCB as a major requirement for a solid architecture.
%A brief comparison of TCB of widespread GPH is depicted in figure \ref{fig:tcbchart}. 
%More importanly, this is particularly true looking at the monolithic approach of commercial hyper-visors, where any addition to the hypervisor body enlarges the TCB.

Finally, the above architecture does not improve the overall stability of the infrastructure. The underlying core, L0, has the same problem of single-layer GPH architecture: the large TCB implies also that a failure experienced in one of the critical components will simply produce a generalized failure of the whole platform, not satysfing DP2. Moreover, in GPH standard architecture, the administrator has the entire control of the platform and could potentially snoop and inspect user VM status and data.

This kind of architecture, in particular, does not leverage completely the possibility to hide behind a compatible layer, a really modular architecture with only a tiny critical core.
To sum up, current nested virtualization limits are represented by a certain lack of dedicated feature in modern hypervisor and the impossibility to rely on a tiny TCB (DP3,DP4).
The first one is being addressed by hypervisor developers. The second one could simply not be achieved with modern hypervisors.

@@

Therefore, leveraging this concept of microkernels, a micro hypervisor (MH) is composed from a small kernel providing only essential functions, like scheduling, IPC, memory management and VMX extension. Several micro-hypervisor have been proposed in literature \cite{xmhf, nova}. They rely on the micro-kernel paradigm, extending it to virtualization. MH naturally satisfy vertical design principles (DP1,DP3, DP4). The tiny core could be considered the only system-critical componente of the architecture and if it possible to assure its correctness it could represent the last line of defense, where it could be possible to rebuild the architecture.

@@

\subsection {MH general issues}

@@

Moreover, as reported in previous sections, several cloud computing issues are simply not addressable with general purpose hypervisor (GPH). However, they are commonly widely adopted and represent the bedrock of existing platform. Therefore, GPH exports wide and solid interfaces leveraged by a plenty of different software that is able to properly communicate and act through them on the virtualization layer. On the contrary, MH lack of cloud toolkit integration, requiring an explicit support to be written, not satysfing 3rd part compatibility requirements (DP5). GPH are tightly connected with cloud toolkit and could present solid interfaces to communicate with different tools.

Finally, for sake of keeping the TCB small, MH could generally lack of several virtualization features, commonly available on modern GPH hypervisor. For instance, XMHF does not support multiple guests. This kind of drawbacks could limits the potential interest in MH, even they are able to introduce several important features.

@@

The first analyzed architecture concerns only GPH. At first, it is important to recall the most interesting achievable by such architectures. In particular, several remarkable features are live migration over different platform, transparent migration of virtualized clusters between different data-centers and improved isolation between customer instances and administration, improved control of the nested virtualization layer L1 (horizontal requirements DP1).
Supposing the completeness and the correctness of nested virtualization hardware facilities in widespread hypervisor, GPH over GPH represents the best achievable architecture to respect horizontal design principles, but it lacks in vertical DPs as the size of TCB is the same of currently deployed architectures.

To sum up, it could be possible to achieve live-migration between different platform and other features related to VM management flexibility but it could not still be possible to reduce the TCB. Analyzing MH approaches, the aim is try at most to retain to features, trying to introduce isolation and privacy improvements due to the lower TCB promised.

@@

MH on MH in reported only to see the orthogonality between this approach and GPH over GPH.
A theoretical implementation of MH over MH will provide the vertical requirements that are missing in the previous approach. The tiny core could be able to overcome monolithic problems due to privilege level of adminitration tasks (DP1), the possibility to improve the resilience of the platform to a large set of software failures (DP2) and finally reduce the possibility of the vulnerabilities of the most privilege part of the software (DP3). However, the interest on such architecture is limited because it is not able to overcome any cited issues of single-layer MH. Moreover, several horizontal properties are theoretically achievable but they will require a huge effort in developing functions of interest. Moreover, the lack of nested VMX functionalities prevents to concretely test this kind of architecture, because MH normally rely on VMX facilities in order to create a nested HVM guests.
To conclude, nested virtualization is exploited in order to fill eco-systems gaps with GPH but MH over MH does not provide any improvement in this sense.

@@

The third architecture, proposed in order to reduce security concerns, is built combining an GPH at L0 and a MH at L1. 
%Several MH are proposed in literature focusing on different .
%Therefore, this model tries to exploit microhypervisor as L1 hypervisor in order to enforce better isolation between L2 and L0. L1 will exploit the emulated nested VMX provided by GPH L0 and virtualize L2. 
%With this approach, we try to isolate the attacker in a VM virtualized on a secured layer, in order to prevent attacks to L0. The attacker would have to face with an essential virtualization layer and it would not be able to attack the back-end virtualization layer. 
An interesting point of this approach is the fact of being easily be deployed, without requiring any extra code addition to the GPH, L0.  Existing GPH-based platforms could potentially exports nested VMX features to those guests. Even if this is not so usual in modern cloud platform, this lack is considered temporary and related only to the experimental status of those features.

However, the amount of code running in L0 kernel space is still as big as before and, therefore, no TCB reduction. 
Finally, adopting a MH as L1 prevents to use inter-platform enhanced features. As analyzed when dealing with MH design, the price of having a tiny core is normally paid with a lack of certain functionalities. For example, XMHF only supports one single guest. It could be easily noticed that this approach is not architecturally able to provide any new "horizontal" features, leveraging on nested virtualization. Therefore, the architecture is only able to provide the same features that are provided by today deployed ones, with an enhanced but not sufficient level of isolation.

@@

\begin{table*}
\label{fin:conf}
\centering
\begin{tabular}{llcllll}
\toprule
Id & Functional design Principle & Class & GPH over GPH & MH over MH & MH over GPH & GPH over MH\\
\midrule
DP1 & User-Data privacy & Vertical & $\times$  & $\checkmark$ & $\times $ & $\checkmark$ \\
      DP2 & Fail-Safe modular architecture & Vertical & $\times$  & $\checkmark$ & $\times$ & $\checkmark$ \\
   DP3 & Small TCB & Vertical &  $\times$& $\checkmark$ & $\times$ & $\checkmark$ \\
   DP4 & Inter-platform support & Horizontal & $\checkmark$  & $\times$ & $\times$ & $\checkmark$ \\
   DP5 & 3rd part compatibility & Horizontal & $\checkmark$  & $\times$ & $\times$ & $\checkmark$ \\
   NDP1 & Minimize code addition & - & $\checkmark$  & $\times$ & $\times$ & $\checkmark$ \\
\bottomrule
      \end{tabular}
\end{table*}

In this configuration, the TCB is smaller than other architectures, due to the modular structure of micro-kernels., executed at L0. Leveraging as L1 a GPH, like Xen or KVM, provides to the user an environment apparently  identical to the one used in production today, satisfying DP5. The L1 GPH is normally integrated with modern cloud toolkits like OpenStack and will support 3rd part applications.
Therefore, the architecture presents a stronger isolation leveraging on nested virtualization, providing in particular a strong tiny core(DP3). The MH layer will be transparent from the user and it will not not any change in the functioning of the platform.

The first class of possible drawbacks of this architecture is represented by the necessity to expose to the L1 nested hypervisor the nested VMX extension to let him ``nest'' another HVM guest. This is experimentally done by GPH like Xen and KVM, but it is generally not included in the essential features provided by a minimal micro-hypervisor. As reported in XMHF paper \cite{xmhf}, the amount of lines of code for nested VMX is around 1KLOC. 
More generally, from an architectural point of view, the general MH minimalist attitude could represent a limitation to general preservation of existing platform functions. The constraint of privileged code size forces normally MH developers to drop some features, considered not essential. Therefore, several limitations could arise not only for nested virtualization proper features, but also for other class of features. For instance, XHMF is able only to virtualize one L1 instance at time. This is acceptable if the platform objective is limited to be provided of a secure core, but dramatically limits the possibility of having a platform model able to give the user the possibility to run its proper hypervisor as L1.

@@

To conclude, this section provides a short comparison between different platforms, contrasting architecture features above the two main different types of benefits given by nested virtualization, as reported widely in literature \cite{art:blan, nested:berghmans,  cloudvisor:zhang, turtle:ibm}. The first could be identified in the augmented degree of isolation the architecture could enforce between host OS and the "customer" point of access, allowing to redesign the security architecture of the whole platform. The second one is represented by the possibility to leverage multiple virtualization layers to easily define inter-platform policy and operations between different platforms.
In table \ref{fin:conf}, it is possible a summary of the previous analysis. To conclude, GPH over MH is the most interesting one in order to enlarge nested virtualization promised benefit. This architecture with a solid and secured MH as L0 is the only able to reduce the platform TCB and this could be a key point to obtain more protected infrastructure, .
Firstly, adopting a MH as architecture trusted core, we will have a secure layer guaranteeing that administrator could not lose the control of the system. Klein et Al. \cite{Klein:2009} have demonstrated that a kernel with less than 10KLoC could be formally verified, under several conditions. This is far from be achievable today with standard GPH over GPH approach where the core is not tiny and the TCB is an order of 
magnitude bigger than MH \cite{nova}.

However, the effort to set up a GPH over MH is strongly dependent on which technique leveraging for virtualizing L2. As stated in paragraph \ref{par:gom}, with L2 guest virtualized with the hardware assitance technique, the most popular one, the MH needs to export nVMX at least. 
On the contrary, adopting paravirtualization/Dynamic binary translation for L2 does not necessitate a particular hardware support and so it could not be necessary patch the MH to set up a platform. 
However, it has not be underestimated, however, that para-virtualization supports only a subset of OS, and could not be considered a general solution and DBT has shown important overhead.
%The last point of interest is related to device management. If MH act as L0 hypervisor, it has to deal with hardware devices, and this could necessitate of a significant code writing. In fact, there are several technologies, like Direct Assignment, allowing to bypass hypervisor control of a device. Those techniques are useful but they reduce the flexibility in the device management, requiring for example exclusive allocation of a device to a guest. To contrast this drawbacks, PCI-E propose the possibity to multiplex device with a set of virtual devices called virtual functions proposed to the L0 hypervisor, that will relax the contraint of exclusive allocation. Unfortunately this approach could not give the same flexibilty given by other techniques, like para-virtualization 
%With MH over GPH, MH could directly assigne devices to GPH that is able to implement  para-virtualize device, for example. 
%This will relief the MH of a part of devices management, giving to the GPH the responsibility to drive concretely the device but without losing interesting features. Therefore, we will need a limited amount of code to handle device without have to deal with each specific device.
